|

### 2021-08-01 prova_knit_finale.Rmd
belli rega APFoptfun funziona!

\begin{algorithm}APF with optimal transition kernel for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1}^{(i)}g(x_{t-1}) $
\item $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}^{(I_{k})}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1}^{(I_{k})})}-2\big],\tau^{2})$ for $i=1,...N$
\item Set  $\tilde{w}^{(i)}_{t} = N^{-1}$ for $i=1,...,N$
\end{enumerate}
\item otherwise
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2}) \ \ i=1,...,N$ for $i=1,...N$
\item Set $\tilde{w}^{(i)}_{t}=w^{(i)}_{t-1}g(x_{t-1})$ for $i=1,...,N$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

@@

```{r include=FALSE}
# Simulating the data
# -------------------
set.seed(12345)
n     =  50
alpha = -0.1
beta  =  0.9
tau2  =  0.01
tau   = sqrt(tau2)
y     = rep(0,n)
x     = rep(0,n)
x[1]  = alpha/(1-beta)
y[1]  = rnorm(1,0,exp(x[1]))
for (t in 2:n){
  x[t] = alpha + beta*x[t-1] + rnorm(1,0,tau)
  y[t] = rnorm(1,0,exp(x[t]))
}
alpha.true = alpha
beta.true  = beta
tau2.true  = tau2

@@

@@

```{r include=FALSE}
N <- 5000
alpha <- -0.1
beta  <-  0.9
tau2  =  0.01
tau   = sqrt(tau2)
m0      = 0
C0      = 100
ealpha  = alpha
valpha  = 0.01
ebeta    = beta
vbeta    = 0.01
nu      = 3
lambda  = tau2

@@

svapf<-SVAPFoptfun(y,N,m0,C0,alpha,beta=0,tau)

@@

```{r echo=FALSE, fig.align='center', fig.cap="Comparison of filtering strategies for Stochastic Volatility Model. Filtered States with credible interval (in red)", fig.height=6, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
plotone<-SVFiltplot(dfsv,svpf,"",1.5,-2.5)
plottwo<-SVFiltplot(dfsv,svgpf,"",1.5,-2.5)
plotthree<-SVFiltplot(dfsv,svapf,"",1.5,-2.5)
plotfour<-SVFiltplot(dfsv,svlw,"",1.5,-2.5)

@@

```{r}
set.seed(123)
truex<-x
estimatedx<-matrix(NA,ncol=5,nrow=4)
colnames(estimatedx)<-c("N","BPF","GPF", "APF","LWF")
estimatedx[,1]<-c(100,1000,10000,100000)

>>>>

\begin{algorithm}APF with optimal transition kernel for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1}^{(i)}g(x_{t-1}) $
\item $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}^{(I_{k})}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1}^{(I_{k})})}-2\big],\tau^{2})$ for $i=1,...N$
\item Set  $\tilde{w}^{(i)}_{t} = N^{-1}$ for $i=1,...,N$
\end{enumerate}
\item otherwise
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2}) \ \ i=1,...,N$ for $i=1,...N$
\item Set $\tilde{w}_{t}^{(i)} = w_{t-1}^{(i)}\frac{f_{N}(y_{t};0,e^{x_{t}})f_{N}(x_{t};\alpha+\beta x_{t-1},\tau^{2})}{f_{N}(x_{t};\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2})}$ for $i=1,...,N$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

@@

```{r include=FALSE}
# Simulating the data
# -------------------
set.seed(12345)
n     =  50
alpha =  0
beta  =  0.9
tau2  =  0.01
tau   = sqrt(tau2)
y     = rep(0,n)
x     = rep(0,n)
x[1]  = alpha/(1-beta)
y[1]  = rnorm(1,0,exp(x[1]))
for (t in 2:n){
  x[t] = alpha + beta*x[t-1] + rnorm(1,0,tau)
  y[t] = rnorm(1,0,exp(x[t]))
}
alpha.true = alpha
beta.true  = beta
tau2.true  = tau2

@@

      ESS  = 1/sum(w^2)

    if(ESS<(N/r)){
      epsilon=alpha+beta*x+((tau^2)/4)*((data[t]^2)*exp(-alpha-beta*x)-2)
    weight = w*exp((1/(2*tau^2))*((epsilon)^2-(alpha+beta*x)^2)-((data[t]^2)/2)*exp(-alpha-beta*x)*(1+alpha+beta*x))
    k   = sample(1:N,size=N,replace=TRUE,prob=weight)
    x1   = rnorm(N,alpha+beta*x[k]+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*x[k]))-2),tau)
    w   = rep(1/N,N)
    }else{
    weight = rep(1/N,N)
    k   = sample(1:N,size=N,replace=FALSE,prob=weight)
    x1   = rnorm(N,alpha+beta*x[k]+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*x[k]))-2),tau)
    w   <-  w*dnorm(data[t],0,exp(x1))*dnorm(x1,alpha+beta*x[k],tau)/
      dnorm(x1,mean=alpha+beta*x[k]+(tau^2)/4*((data[t]^2)*exp(-(alpha+beta*x[k]))-2),tau)
    }

@@

```{r include=FALSE}
N <- 5000
alpha <- 0
beta  <-  0.9
tau2  =  0.01
tau   = sqrt(tau2)
m0      = 0
C0      = 100
ealpha  = alpha
valpha  = 0.01
ebeta    = beta
vbeta    = 0.01
nu      = 3
lambda  = tau2

@@

svapf<-SVAPFoptfun(y,N,m0,C0,alpha,beta,tau)

@@

```{r echo=FALSE, fig.align='center', fig.cap="Comparison of filtering strategies for Stochastic Volatility Model. Filtered States with credible interval (in red)", fig.height=6, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
plotone<-SVFiltplot(dfsv,svpf,"",2,-2)
plottwo<-SVFiltplot(dfsv,svgpf,"",2,-2)
plotthree<-SVFiltplot(dfsv,svapf,"",2,-2)
plotfour<-SVFiltplot(dfsv,svlw,"",2,-2)

@@

```{r}
set.seed(123)
truex<-x
estimatedx<-matrix(NA,ncol=5,nrow=5)
colnames(estimatedx)<-c("N","BPF","GPF", "APF","LWF")
estimatedx[,1]<-c(10,100,1000,10000,100000)
